Bioverse: Origins of Life

### 2023-08-11 src/tex/ms.tex
remove ORCID to debug syw
showyourwork seems to complain when my ORCID is included in \author:
"
Preprocess: Setting up the workflow...
Traceback (most recent call last):
  File "/Users/schlecker/repos/originsoflife/.snakemake/scripts/tmps6etcd0c.preprocess.py", line 495, in <module>
    config["tree"] = get_json_tree(snakemake.input[0])
  File "/Users/schlecker/repos/originsoflife/.snakemake/scripts/tmps6etcd0c.preprocess.py", line 411, in get_json_tree
    free_floating_graphics = [
  File "/Users/schlecker/repos/originsoflife/.snakemake/scripts/tmps6etcd0c.preprocess.py", line 414, in <list

% Author list
\author[0000-0001-8355-2107]{Martin Schlecker}
\affiliation{Steward Observatory, The University of Arizona, Tucson, AZ 85721, USA; \href{mailto:schlecker@arizona.edu}{schlecker@arizona.edu}}
\author{et al.}

@@

\section{Bayesian Analysis}
It is instructive to consider the constraining power of a successful biosignature detection for competing OoL scenarios, which we here attempt with an analytical approach.
We may first consider the probability $\pdf(bio|H_i)$ of detecting a convincing biosignature on a planet, given a specific OoL hypothesis $H_i, \, i \in 1, 2, 3$ is true.
This can be decomposed into the probability of abiogenesis $\pdf_\mathrm{abio}$, the fraction of life-hosting worlds that develop atmospheric biosignatures $\pdf_\mathrm{sig}$, the probability of the planet type required by the OoL hypothesis occurring in the surveyed sample $\pdf_\mathrm{\eta, i}$, and the probability of detecting the biosignature on this planet type with current technology $\pdf_\mathrm{det, i}$, yielding

@@

However, what we are actually interested in is the probability of a OoL hypothesis being true, given a specific biosignature detection, $\pdf(H_i|bio)$.
To obtain this we can use Bayes' theorem, which yields
\begin{align}
\pdf(H_i|bio) = \frac{\pdf(bio|H_i) \pdf(H_i)}{\pdf(bio)}.
\end{align}
Here, $\pdf(H_i)$ is the prior probability of the OoL hypothesis $H_i$, and $\pdf(bio)$ is the prior probability of detecting a biosignature.
% Following Eq.~\ref{eqn:pbio}, we can sum over all possible scenarios to obtain
% \begin{align}
% \pdf(bio) = \sum_{i=1}^{N} \pdf_\mathrm{abio} \times \pdf_\mathrm{sig} \times \pdf_\mathrm{\eta, i} \times \pdf_\mathrm{det, i}.
% \end{align}

@@

\subsection{Fractional planet occurrence rate $\pdf_\mathrm{\eta, i}$}
As the different OoL hypotheses do not all work on the same planet types, we may study the impact of the fractional occurrence rates of different planet types on the posterior probability $\pdf(H_i|bio)$.
For simplicity, let's consider only planets that allow at least one of the scenarios.
We also ignore any influence of other planets in the same system, e.g. an outer gas giant that itself does not develop life~\citep{Schlecker2021a} or panspermia scenarios CITE.
We may then distinguish between:
\begin{enumerate}
    \item Earths, i.e., limited-water terrestrial planets in the liquid water habitable zone ($H1, H2, H3$). These planets with roughly Earth-like water mass fractions support both the existence of submarine hydrothermal vents ($H1$) and hydrothermal fields with wet/dry cycles ($H2$). Limits in exoplanet observables to this planet type are their orbital distance (both scenarios require liquid water; at least for FGK stars this requirement also puts the planet outside of the tidal locking zone ($H2$)), and bulk density ($H1$, $H2$, and $H3$ require limited water fractions).
    \item ``shallow-water'' water worlds ($H1$). These planets have no land surface exposed to the atmosphere, thus excluding the subaerial pond scenario.
    \item warm terrestrial planets ($H3$). These planets have smaller orbital distances, which serves as a proxy for high past UV fluxes (as required by $H3$).
    \item ``deep ocean'' water worlds. Through the development of high-pressure ices, these planets do not support any of the considered OoL scenarios.
\end{enumerate}

@@

>>>>

% Author list
\author{Martin Schlecker}
%\author[0000-0001-8355-2107]{Martin Schlecker}
\affiliation{Steward Observatory, The University of Arizona, Tucson, AZ 85721, USA; \href{mailto:schlecker@arizona.edu}{schlecker@arizona.edu}}
\author{et al.}

@@

\section{Bayesian Analysis}
\todo[inline]{consider Fisher matrix (Kendall \& Stuart 1977; Tegmark 1997) analysis to determine sensitivity of a survey to a set of parameters?}
It is instructive to consider the constraining power of a successful biosignature detection for competing OoL scenarios, which we here attempt with an analytical approach.
The hydrothermal vents scenario ($H_1$) and the subaereal pond scenario ($H_2$) can be considered as mutually exclusive models, and we may study how a particular future observation of biosignatures impacts our beliefs about the relative model probabilities.

We may first consider the probability $\pdf(bio|H_i)$ of detecting a convincing biosignature on a planet, given a particular OoL hypothesis $H_i, \, i \in 1, 2, 3$ is true.
This can be decomposed into the probability of abiogenesis $\pdf_\mathrm{abio}$, the fraction of life-hosting worlds that develop atmospheric biosignatures $\pdf_\mathrm{sig}$, the probability of the planet type required by the OoL hypothesis occurring in the surveyed sample $\pdf_\mathrm{\eta, i}$, and the probability of detecting the biosignature on this planet type with current technology $\pdf_\mathrm{det, i}$, yielding

@@

However, what we are actually interested in is the probability of a OoL hypothesis being true, given a particular biosignature detection, $\pdf(H_i|bio)$.
To obtain this we can use Bayes' theorem, which yields
\begin{align}
\pdf(H_i|bio) = \frac{\pdf(bio|H_i) \pdf(H_i)}{\pdf(bio)}.
\end{align}
Here, $\pdf(H_i)$ is the prior probability of the OoL hypothesis $H_i$, and $\pdf(bio)$ is the prior probability of detecting a biosignature.
% Following Eq.~\ref{eqn:pbio}, we can sum over all possible scenarios to obtain
% \begin{align}
% \pdf(bio) = \sum_{i=1}^{N} \pdf_\mathrm{abio} \times \pdf_\mathrm{sig} \times \pdf_\mathrm{\eta, i} \times \pdf_\mathrm{det, i}.
% \end{align}

@@

\subsection{Fractional planet occurrence rate $\pdf_\mathrm{\eta, i}$}
As the different OoL hypotheses do not all work on the same planet types, we may study the impact of the fractional occurrence rates of different planet types on the posterior probability $\pdf(H_i|bio)$.
For simplicity, let's consider only planets that allow at least one of the scenarios.
We also ignore any influence of other planets in the same system, e.g. an outer gas giant that itself does not develop life~\citep{Schlecker2021a} or panspermia scenarios CITE.
We may then distinguish between:
\begin{enumerate}
    \item Earths, i.e., limited-water terrestrial planets in the liquid water habitable zone ($H1, H2, H3$). These planets with roughly Earth-like water mass fractions support both the existence of submarine hydrothermal vents ($H1$) and hydrothermal fields with wet/dry cycles ($H2$). Limits in exoplanet observables to this planet type are their orbital distance (both scenarios require liquid water; at least for FGK stars this requirement also puts the planet outside of the tidal locking zone ($H2$)), and bulk density ($H1$, $H2$, and $H3$ require limited water fractions).
    \item ``shallow ocean'' water worlds ($H1$). These planets have no land surface exposed to the atmosphere, thus excluding the subaerial pond scenario.
    \item warm terrestrial planets ($H3$). These planets have smaller orbital distances, which serves as a proxy for high past UV fluxes (as required by $H3$).
    \item ``deep ocean'' water worlds. Through the development of high-pressure ices, these planets do not support any of the considered OoL scenarios.
\end{enumerate}

@@

\subsection{Discussion of Bayesian analysis}

\begin{note}
    shallow ocean planets are vulnerable to water loss through high-energy radiation, limiting the time window for habitability especially if no geochemical feedbacks exist~\citep{Kite2018}.
\end{note}
\todo[inline]{discuss host star spectral type dependencies, e.g., abiogenesis time window for G stars ($\lessapprox 10 Gyr$, MS lifetime) or M dwarfs. \citep[e.g.,][]{Spiegel2012}}
